---
title: "4. Results"
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    fig_caption: TRUE
    number_sections: FALSE
    global_numbering: FALSE
    pandoc_args: !expr acronymsdown::add_filter()
  # pdf_document:
  #   latex_engine: xelatex
  #   dev: cairo_pdf
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                        encoding = encoding,
                        knit_root_dir = rprojroot::find_rstudio_root_file(),
                        output_dir = file.path("../Output")
                        )})
bibliography: ../Script/Lit_Rev_Refs.bib
csl: ../Script/vancouver-superscript.csl
link-citations: TRUE
urlcolor: blue  
acronyms:
  loa_title: ""
  include_unused: TRUE
  insert_loa: "end"
  insert_links: TRUE
  id_prefix: "acronyms_"
  sorting: "alphabetical"
  non_existing: "key"
  style: "long-short"
  fromfile: ./_acronyms.yml
  
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE, 
                      message = FALSE, 
                      error = FALSE,
                      cache = FALSE, 
                      include = FALSE, 
                      dev = "png", 
                      dpi = 500
)

# Load packages
library(tidyverse)
library(dplyr)
library(bookdown)
library(TwoSampleMR)
library(rstan)
library(here)
library(cowplot)
library(kableExtra)
library(tables)
library(flextable)
library(ftExtra)
library(officer)

# Load University of Edinburgh colour palette
source(here("MSc_Thesis_Split", "Script", "edin_uni_colours.R"))

# Load pre-formatted plot template - call to ggplot with UoE colours
source(here("MSc_Thesis_Split", "Script", "edin_fig_style.R"))

# Load function scripts
source(here("MSc_Thesis_Split", "Script", "simulation_functions.R"))

# Compile model for MR-Hevo
mr.stanmodel <- stan_model(file = here("MSc_Thesis_Split",
                                       "Script",
                                       "Hevo",
                                       "MRHevo_summarystats.stan"),
                           model_name = "MRHevo.summarystats", 
                           verbose = FALSE)


# --- Load data --- #
## 0% Invalid
no_causal_10k_point0_scen1_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "no_causal_10k_point0_scen1_models.rds"))
causal_10k_point0_scen1_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "causal_10k_point0_scen1_models.rds"))

## 10% Invalid
no_causal_10k_point1_scen1_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "no_causal_10k_point1_scen1_models.rds"))
causal_10k_point1_scen1_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "causal_10k_point1_scen1_models.rds"))

## 20% Invalid
no_causal_10k_point2_scen2_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "no_causal_10k_point2_scen2_models.rds"))
causal_10k_point2_scen2_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "causal_10k_point2_scen2_models.rds"))

## 30% Invalid
no_causal_10k_point3_scen3_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "no_causal_10k_point3_scen3_models.rds"))
causal_10k_point3_scen3_models <- readRDS(file = here("MSc_Thesis_Split", "Data", "Simulated_Datasets", "causal_10k_point3_scen3_models.rds"))


# --- Load results --- #
# No Causal:
no_causal_10k_scen1_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "no_causal_10k_scen1_sim_summ_tib.rds"))
#no_causal_20k_scen1_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "no_causal_20k_scen1_sim_summ_tib.rds"))

no_causal_10k_scen2_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "no_causal_10k_scen2_sim_summ_tib.rds"))
#no_causal_20k_scen2_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "no_causal_20k_scen2_sim_summ_tib.rds"))

no_causal_10k_scen3_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "no_causal_10k_scen3_sim_summ_tib.rds"))
#no_causal_20k_scen3_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "no_causal_20k_scen3_sim_summ_tib.rds"))


# Causal:
causal_10k_scen1_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "causal_10k_scen1_sim_summ_tib.rds"))
#causal_20k_scen1_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "causal_20k_scen1_sim_summ_tib.rds"))

causal_10k_scen2_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "causal_10k_scen2_sim_summ_tib.rds"))
#causal_20k_scen2_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "causal_20k_scen2_sim_summ_tib.rds"))

causal_10k_scen3_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "causal_10k_scen3_sim_summ_tib.rds"))
#causal_20k_scen3_sim_summ_tib <- readRDS(file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "causal_20k_scen3_sim_summ_tib.rds"))

# --- Citation Search --- # 

Citations_Instrument_Data <- read.csv(here("MSc_Thesis_Split", "Data", "Citations_Datasets", "Citations_Instrument_Data.csv")) %>% as_tibble() #%>% filter(Instrument !="rs7326482") %>% mutate(N=1)
# Citations_Instrument_Data_2 <- read.csv(here("MSc_Thesis_Split", "Data", "Citations_Datasets", "Citations_Instrument_Data.csv")) %>% as.tibble() #%>%filter(Instrument !="rs7326482") %>%  mutate(N=2)

```

```{r sim-summ-tibs}

# --- No Causal --- #
# Scenario 1
no_causal_scen1_sim_summ_tib <- bind_rows(no_causal_10k_scen1_sim_summ_tib#,
                                          #no_causal_20k_scen1_sim_summ_tib
) %>%
  mutate(Scenario = "Scenario 1: Balanced pleiotropy, InSIDE assumption satisfied")

# Scenario 2
no_causal_scen2_sim_summ_tib <- bind_rows(no_causal_10k_scen2_sim_summ_tib#,
#no_causal_20k_scen2_sim_summ_tib
                                          ) %>%
  mutate(Scenario = "Scenario 2: Directional pleiotropy, InSIDE assumption satisfied")

# Scenario 3
no_causal_scen3_sim_summ_tib <- bind_rows(no_causal_10k_scen3_sim_summ_tib#,
#                                           no_causal_20k_scen3_sim_summ_tib
                                           ) %>%
   mutate(Scenario = "Scenario 3: Directional pleiotropy, InSIDE assumption not satisfied")


# Summary
no_causal_sim_summ_tib <- bind_rows(no_causal_scen1_sim_summ_tib,
                                    no_causal_scen2_sim_summ_tib,
                                    no_causal_scen3_sim_summ_tib
)%>% 
  mutate(WME_Lower_CI = WME_Av - 1.96 * WME_SE,
         WME_Upper_CI = WME_Av + 1.96 * WME_SE,
         Pos_Rate_Diff = Hevo_Pos_Rate - WME_Pos_Rate,
         Hevo_Better = Pos_Rate_Diff < 0)


# --- Causal --- #
# Scenario 1
causal_scen1_sim_summ_tib <- bind_rows(causal_10k_scen1_sim_summ_tib#,
                                       #causal_20k_scen1_sim_summ_tib
) %>% 
  mutate(Scenario = "Scenario 1: Balanced pleiotropy, InSIDE assumption satisfied")

# Scenario 2
causal_scen2_sim_summ_tib <- bind_rows(causal_10k_scen2_sim_summ_tib#,
                                       #causal_20k_scen2_sim_summ_tib
) %>% 
  mutate(Scenario = "Scenario 2: Directional pleiotropy, InSIDE assumption satisfied")

# Scenario 3
causal_scen3_sim_summ_tib <- bind_rows(causal_10k_scen3_sim_summ_tib#,
                                       #causal_20k_scen3_sim_summ_tib
) %>% 
  mutate(Scenario = "Scenario 3: Directional pleiotropy, InSIDE assumption not satisfied")


# Summary
causal_sim_summ_tib <- bind_rows(causal_scen1_sim_summ_tib,
                                 causal_scen2_sim_summ_tib,
                                 causal_scen3_sim_summ_tib) %>% 
  mutate(WME_Lower_CI = WME_Av - 1.96 * WME_SE,
         WME_Upper_CI = WME_Av + 1.96 * WME_SE,
         Pos_Rate_Diff = Hevo_Pos_Rate - WME_Pos_Rate,
         Hevo_Better = Pos_Rate_Diff < 0)
  
#no_causal_sim_summ_tib
```

Word count: `r wordcountaddin::word_count(here::here("MSc_Thesis_Split", "Script", "4_Results.Rmd"))`

# Results

## Simulation Study

### Data Simulation

Data were successfully simulated as intended. A selection of representative visualisations are presented in Figure \@ref(fig:data-sim-validation). Full details of testing used to validate model outputs from parameter inputs are given in Appendix \@ref(appendix-sim-test). The $F$-statistic calculated from simulated instruments was >10, indicating that they were sufficiently strongly associated with exposure to meet the relevance assumption of \acr{IV} analysis (Tables \@ref(tab:no-causal-sim-summ-display) and \@ref(tab:causal-sim-summ-display)).

\newpage


```{r data-sim-validation, include=TRUE, fig.height=7.5}
#| fig.id = "data-sim-validation",
#| fig.cap = "Plots of a representative group of simulated datasets; all simulate genetic instruments from the same index from the same random seed. Left and right columns demonstrate null and positive true causal effects, respectively. The scenario and the proportion of invalid (i.e. pleiotropic) genetic instruments changes with each row. a) 10% of instruments invalid, Scenario 1: balanced pleiotropy introduces noise around the causal effect. b) 20% of instruments invalid, Scenario 2: directional pleiotropy biases in the direction of the invalid instruments. c) 30% of instruments invalid, Scenario 3: directional pleiotropy and InSIDE assumption violation strongly biases towards a positive effect estimate."

# Get random dataset index
set.seed(1701)
rand_example <- sample(1:1000, 1, replace = FALSE)

# Select random dataset from each scenario
models_list <- list(
  # 0% Invalid
  no_causal_10k_point0_scen1_models[[rand_example]],
  causal_10k_point0_scen1_models[[rand_example]],
  ## 10% Invalid
  no_causal_10k_point1_scen1_models[[rand_example]],
  causal_10k_point1_scen1_models[[rand_example]],
  ## 20% Invalid
  no_causal_10k_point2_scen2_models[[rand_example]],
  causal_10k_point2_scen2_models[[rand_example]],
  ## 30% Invalid
  no_causal_10k_point3_scen3_models[[rand_example]],
  causal_10k_point3_scen3_models[[rand_example]]
)

# Create blank list to receive plots
Gamma_gamma_plots <- as.list(rep(NA, length(models_list)))

# Create plot subtitles
subtitles <- c(
  "0% Invalid",
  "",
  "10% Invalid, Scenario 1",
  "",
  "20% Invalid, Scenario 2",
  "",
  "30% Invalid, Scenario 3",
  ""
)


# Create plots
for(model in 1:length(models_list)){
  
  # Set beta for graph slope
  true_beta <- if_else(model%%2 == 1,
                       true = 0,
                       false = 0.1) 
  
  # Label only left/bottom axes
  y_lab <- if_else(model%%2 == 1,
                   true = "Gene-Outcome\nCoefficient (\U1D26)",
                   false = "") 
  
  x_lab <- case_when(model == length(models_list) - 1 | model == length(models_list) ~ "Gene-Exposure Coefficient (\U03B3)",
                     .default = "") 
  
  # Title Causal/Non-Causal
  causal_title <- case_when(model ==  1 ~ "No Causal Effect",
                            model == 2 ~ "Causal Effect",
                            .default = "") 
  
  # Tag index
  tag_index <- if_else(model%%2 == 1,
                       true = (model+1)/2,
                       false = FALSE) 
  
  # Tag text
  tag_text <- if_else(model%%2 == 1,
                       true = paste0(letters[tag_index], ")"),
                       false = "") 
  
  
  # Plot outline
  Gamma_gamma_plots[[model]] <- models_list[[model]] %>%
    tibble() %>% 
    mutate(Invalid = (alpha != 0)) %>%
    plot_template() +
    aes(x = coeff_G_X,
        y = coeff_G_Y) +
    # Instruments
    geom_point(aes(fill = Invalid),
               pch = 21,
               alpha = 0.75,
               size = 3) +
    # Lines
    geom_abline(colour = edin_uni_blue_hex,
                slope = true_beta,
                intercept = 0,
                size = 1) +
    
    # Add betas
    ## True
    geom_text(label = paste0("True \U03B2 = ", true_beta), #beta
              x = 0.1, # labels with gradient (causal effect estimate)
              y = -0.1,
              size = 3,
              fontface = "italic",
              colour = edin_uni_blue_hex,
              hjust = 0,
              data = . %>% slice_head() # prevent over-printing
    ) +
    # Colours for valid/invalid instruments
    scale_fill_manual(values = c(edin_muted_turquoise_hex, edin_burgundy_hex),
                      labels = c("Valid", "Invalid"),
                      guide = "legend",
                      name = "Instruments") +
    # Labels
    labs(#title = causal_title,
         subtitle = subtitles[model],
         tag = tag_text,
         x = x_lab,
         y = y_lab) +
    theme(legend.position = "none",
          plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"),
          axis.text = element_text(size = 6), 
          axis.title = element_text(size = 7), 
          subtitle = element_text(size = 6)
          ) +
    xlim(-0.06, 0.2125) +
    ylim(-0.11, 0.25) 
  
}

#Legend
legend <- get_legend(Gamma_gamma_plots[[length(models_list)]] +
                       theme(legend.position = "right")
                     )

plot_two_cols <- plot_grid(plotlist = Gamma_gamma_plots, 
                           rel_widths = c(0.55, 0.45),
                           ncol = 2)


# Plot all
plot_grid(plot_two_cols,
          legend,
          ncol = 2,
          rel_widths = c(1, 0.2))

```


\newpage

### Analysis of Simulated Data

```{r inline-vals-no-causal}

# --- No Causal --- #
# Mean positive report rates
no_causal_Hevo_mean_pos_rate <- mean(no_causal_sim_summ_tib$Hevo_Pos_Rate * 100) %>% signif(2)
no_causal_WME_mean_pos_rate <- mean(no_causal_sim_summ_tib$WME_Pos_Rate * 100) %>% signif(2)

# Proportion of combinations of scenarios/parameters
no_causal_comb_n <- no_causal_sim_summ_tib$Hevo_Better %>% length()
no_causal_Hevo_better_n <- no_causal_sim_summ_tib$Hevo_Better %>% sum(na.rm = TRUE)
no_causal_Hevo_better_pc <- (no_causal_Hevo_better_n/no_causal_comb_n * 100) %>% signif(2)

# Mean/CIs
no_causal_Hevo_mean_est <- no_causal_sim_summ_tib$Hevo_Av %>% mean() %>% signif(2)
no_causal_Hevo_mean_lower <- no_causal_sim_summ_tib$Hevo_Lower_CI %>% mean() %>% signif(2)
no_causal_Hevo_mean_upper <- no_causal_sim_summ_tib$Hevo_Upper_CI %>% mean() %>% signif(2)

no_causal_WME_mean_est <- no_causal_sim_summ_tib$WME_Av %>% mean() %>% signif(2)
no_causal_WME_mean_lower <- no_causal_sim_summ_tib$WME_Lower_CI %>% mean() %>% signif(2)
no_causal_WME_mean_upper <- no_causal_sim_summ_tib$WME_Upper_CI %>% mean() %>% signif(2)


# SEs
no_causal_Hevo_mean_SE <- no_causal_sim_summ_tib$Hevo_SE %>% mean() %>% signif(2)
no_causal_Hevo_min_SE <- no_causal_sim_summ_tib$Hevo_SE %>% min() %>% signif(2)
no_causal_Hevo_max_SE <- no_causal_sim_summ_tib$Hevo_SE %>% max() %>% signif(2)

no_causal_WME_mean_SE <- no_causal_sim_summ_tib$WME_SE %>% mean() %>% signif(2)
no_causal_WME_min_SE <- no_causal_sim_summ_tib$WME_SE %>% min() %>% signif(2)
no_causal_WME_max_SE <- no_causal_sim_summ_tib$WME_SE %>% max() %>% signif(2)

```

#### No Causal Effect
\leavevmode\newline Across all cases where no causal effect was present (Table \@ref(tab:no-causal-sim-summ-display)), the mean rate of reporting a causal effect (i.e. false positive rate) for MR-Hevo was `r no_causal_Hevo_mean_pos_rate`% versus `r no_causal_WME_mean_pos_rate`% for \acr{WME}. Of the `r no_causal_comb_n` combinations of scenarios and parameters, MR-Hevo exhibited a favourable false positive rate versus \acr{WME} in `r no_causal_Hevo_better_n` (`r no_causal_Hevo_better_pc`%). For both MR-Hevo and \acr{WME} methods, causal report rate generally increased with an increasing proportion of invalid instruments, and also as assumption violations progressively presented greater data variability and bias across scenarios 1 to 3.

The mean causal effect estimate (mean reported 95% \acr{CI}s) across all cases was `r no_causal_Hevo_mean_est` (`r no_causal_Hevo_mean_lower` to `r no_causal_Hevo_mean_upper`) for MR-Hevo and `r no_causal_WME_mean_est` (`r no_causal_WME_mean_lower` to `r no_causal_WME_mean_upper`) for \acr{WME}. The mean (range) \acr{SE} of causal effect estimates across all cases was `r no_causal_Hevo_mean_SE` (`r no_causal_Hevo_min_SE` to `r no_causal_Hevo_max_SE`) for MR-Hevo and `r no_causal_WME_mean_SE` (`r no_causal_WME_min_SE` to `r no_causal_WME_max_SE`). For both MR-Hevo and \acr{WME} methods, causal effect estimates, width of \acr{CI}s and \acr{SE} all tended to increase slightly both with an increasing proportion of invalid instruments, and as assumption violations progressively presented greater data variability and bias across scenarios 1 to 3.

```{r inline-vals-causal}

# --- No Causal --- #
# Mean positive report rates
causal_Hevo_mean_pos_rate <- mean(causal_sim_summ_tib$Hevo_Pos_Rate * 100) %>% signif(2)
causal_WME_mean_pos_rate <- mean(causal_sim_summ_tib$WME_Pos_Rate * 100) %>% signif(2)

# Proportion of combinations of scenarios/parameters
causal_comb_n <- causal_sim_summ_tib$Hevo_Better %>% length()
causal_Hevo_better_n <- causal_sim_summ_tib$Hevo_Better %>% sum(na.rm = TRUE)
causal_Hevo_better_pc <- (causal_Hevo_better_n/causal_comb_n * 100) %>% signif(2)

# Mean/CIs
causal_Hevo_mean_est <- causal_sim_summ_tib$Hevo_Av %>% mean() %>% signif(2)
causal_Hevo_mean_lower <- causal_sim_summ_tib$Hevo_Lower_CI %>% mean() %>% signif(2)
causal_Hevo_mean_upper <- causal_sim_summ_tib$Hevo_Upper_CI %>% mean() %>% signif(2)

causal_WME_mean_est <- causal_sim_summ_tib$WME_Av %>% mean() %>% signif(2)
causal_WME_mean_lower <- causal_sim_summ_tib$WME_Lower_CI %>% mean() %>% signif(2)
causal_WME_mean_upper <- causal_sim_summ_tib$WME_Upper_CI %>% mean() %>% signif(2)


# SEs
causal_Hevo_mean_SE <- causal_sim_summ_tib$Hevo_SE %>% mean() %>% signif(2)
causal_Hevo_min_SE <- causal_sim_summ_tib$Hevo_SE %>% min() %>% signif(2)
causal_Hevo_max_SE <- causal_sim_summ_tib$Hevo_SE %>% max() %>% signif(2)

causal_WME_mean_SE <- causal_sim_summ_tib$WME_SE %>% mean() %>% signif(2)
causal_WME_min_SE <- causal_sim_summ_tib$WME_SE %>% min() %>% signif(2)
causal_WME_max_SE <- causal_sim_summ_tib$WME_SE %>% max() %>% signif(2)

```

#### Positive Causal Effect
\leavevmode\newline Across all cases where positive causal effect was present (Table \@ref(tab:causal-sim-summ-display)), the mean rate of reporting a causal effect (i.e. sensitivity) for MR-Hevo was `r causal_Hevo_mean_pos_rate`% versus `r causal_WME_mean_pos_rate`% for \acr{WME}. Of the `r causal_comb_n` combinations of scenarios and parameters, MR-Hevo exhibited a favourable sensitivity versus \acr{WME} in `r causal_Hevo_better_n` (`r causal_Hevo_better_pc`%). For both MR-Hevo and \acr{WME} methods, causal report rate generally increased with an increasing proportion of invalid instruments, and also as assumption violations progressively presented greater data variability and bias across scenarios 1 to 3.

The mean causal effect estimate (mean reported 95% \acr{CI}s) across all cases was `r causal_Hevo_mean_est` (`r causal_Hevo_mean_lower` to `r causal_Hevo_mean_upper`) for MR-Hevo and `r causal_WME_mean_est` (`r causal_WME_mean_lower` to `r causal_WME_mean_upper`) for \acr{WME}. The mean (range) \acr{SE} of causal effect estimates across all cases was `r causal_Hevo_mean_SE` (`r causal_Hevo_min_SE` to `r causal_Hevo_max_SE`) for MR-Hevo and `r causal_WME_mean_SE` (`r causal_WME_min_SE` to `r causal_WME_max_SE`). For both MR-Hevo and \acr{WME} methods, causal effect estimates, width of \acr{CI}s and \acr{SE} all tended to increase slightly both with an increasing proportion of invalid instruments, and as assumption violations progressively presented greater data variability and bias across scenarios 1 to 3.

\newpage

## Tables

```{r no-causal-sim-summ-display, echo=FALSE, warning=FALSE, cache=TRUE, include=TRUE, tab.id="no-causal-sim-summ-display", tab.cap="Summary of 1,000 simulated Mendelian randomisation studies per combination of scenario and parameters, all with null causal effect", results='asis'}
#| ft.arraystretch = 1


# Display
no_causal_sim_summ_tib_display <- no_causal_sim_summ_tib %>%
  # Change names for cleaner plotting
  mutate(Invalid = paste0((Prop_Invalid * 100), "%"),
         F = round(F_stat, 1),
         R2 = paste0(R2_stat *100, "%"),
         Weighted_Median_Mean_Estimate = WME_Av,
         Weighted_Median_Mean_SE = paste0("(", WME_SE, ")"),
         Weighted_Median_Lower_CI = round(WME_Av - 1.96 * WME_SE, 2),
         Weighted_Median_Upper_CI = round(WME_Av + 1.96 * WME_SE, 2),
         Weighted_Median_95_CI = paste0(Weighted_Median_Lower_CI, " to ", Weighted_Median_Upper_CI),
         Weighted_Median_Causal_Report_Rate = paste0((WME_Pos_Rate * 100), "%"),
         MR_Hevo_Mean_Estimate = Hevo_Av,
         MR_Hevo_Mean_SE = paste0("(", Hevo_SE, ")"),
         MR_Hevo_Lower_CI = round(Hevo_Lower_CI, 2),
         MR_Hevo_Upper_CI = round(Hevo_Upper_CI, 2),
         MR_Hevo_95_CI = paste0(round(Hevo_Lower_CI, 2), " to ", round(Hevo_Upper_CI, 2)),
         MR_Hevo_Causal_Report_Rate = paste0((Hevo_Pos_Rate * 100), "%")) %>%
  # Drop old names
  select(N,
         Invalid,
         F,
         R2,
         Weighted_Median_Mean_Estimate,
         Weighted_Median_Mean_SE,
         Weighted_Median_95_CI,
         Weighted_Median_Causal_Report_Rate,
         MR_Hevo_Mean_Estimate,
         MR_Hevo_Mean_SE,
         MR_Hevo_95_CI,
         MR_Hevo_Causal_Report_Rate,
         Scenario) %>%
  as_grouped_data(groups = c("Scenario")) %>%
  as_flextable(hide_grouplabel = TRUE, col_keys = c("N",
                                                    "Invalid",
                                                    "F",
                                                    "R2",
                                                    "Weighted_Median_Mean_Estimate",
                                                    "Weighted_Median_Mean_SE",
                                                    "Weighted_Median_95_CI", 
                                                    "Weighted_Median_Causal_Report_Rate",
                                                    "blank_col",
                                                    "MR_Hevo_Mean_Estimate",
                                                    "MR_Hevo_Mean_SE",
                                                    "MR_Hevo_95_CI", 
                                                    "MR_Hevo_Causal_Report_Rate",
                                                    "Scenario")) |>
  # To separate WME/MR-Hevo
  width(j = "blank_col", width = 0.2) |>
  empty_blanks() |>
  span_header() |>
  align(part = "all", align = "center") |>
  align(part = "body", j =c(5, 9),  align = "right") |>
  align(part = "body", j =c(6, 10),  align = "left") |>
  # Italicise N and F
  italic(part = "header", i = 1, j = c(1,3) ) |>
  # Fix name of invalid IVs
  compose(part = "header", i = 1:4, j = 2, value = as_paragraph("Invalid IVs")) |>
  # Fix name of 95% CIs
  compose(part = "header", i = 3, j = c(7, 12), value = as_paragraph("Mean")) |>
  compose(part = "header", i = 4, j = c(7, 12), value = as_paragraph("95% CI")) |>
  # Fix split headers
  compose(part = "header", i = 3, j = c(5,10), value = as_paragraph("Mean Estimate")) |>
  compose(part = "header", i = 4, j = c(5:6,10:11), value = as_paragraph("(Mean SE)")) |>
  merge_h(part = "header", i = 4) |>
  # Italicise/superscript R^2
  compose(part = "header", i = 1, j = 4, value = as_paragraph(as_i("R"), as_sup(as_i(as.character(2))))) |>
  
  hline(part = "header", i = 2, j = c(5:8, 10:13), border = fp_border(color = "black", width = 1)) |>
  #footnote(i = c(1,4), j = c(2, 5), part = "header", value = as_paragraph(c("IV: Instrumental Variable", "SE: Standard Error")))
  # Fix dimensions
  autofit() |>
  #line_spacing(space = 0, part = "header") |>
  line_spacing(space = 1, part = "body") |>
  fit_to_width(max_width = 6.5) |>
  #height_all(height = 40, unit = "mm",part = "all") |>

  # Add footer
  add_footer_lines("CI: Confidence Interval, IV: Instumental Variable, SE: Standard Error. Null Causal Effect (\U03B2 = 0)") |>
  fontsize(8, i = c(1:5), j = c(1:13), part = "header") |>
  fontsize(8, i = c(1:15), j = c(1:13), part = "body") |>
  fontsize(8, i = 1, j = 1, part = "footer")
# set_caption(caption = "Summary of 1000 Simulated Mendelian Randomisation Studies With Null Causal Effect (\U03B2 = 0)",
#             style = "Table Caption",
#             autonum = run_autonum(seq_id = "tab",
#                                   bkm = "no-causal-sim-summ-tib"))

no_causal_sim_summ_tib_display


```


\newpage
```{r causal-sim-summ-display, echo=FALSE, warning=FALSE, cache=TRUE, include=TRUE, tab.id="causal-sim-summ-display", tab.cap="Summary of 1,000 simulated Mendelian randomisation studies per combination of scenario and parameters, all with positive causal effect", results='asis'}
#| ft.arraystretch = 1


# Display
causal_sim_summ_tib_display <- causal_sim_summ_tib %>%
# Change names for cleaner plotting
  mutate(Invalid = paste0((Prop_Invalid * 100), "%"),
         F = round(F_stat, 1),
         R2 = paste0(R2_stat *100, "%"),
         Weighted_Median_Mean_Estimate = WME_Av,
         Weighted_Median_Mean_SE = paste0("(", WME_SE, ")"),
         Weighted_Median_Lower_CI = round(WME_Av - 1.96 * WME_SE, 2),
         Weighted_Median_Upper_CI = round(WME_Av + 1.96 * WME_SE, 2),
         Weighted_Median_95_CI = paste0(Weighted_Median_Lower_CI, " to ", Weighted_Median_Upper_CI),
         Weighted_Median_Causal_Report_Rate = paste0((WME_Pos_Rate * 100), "%"),
         MR_Hevo_Mean_Estimate = Hevo_Av,
         MR_Hevo_Mean_SE = paste0("(", Hevo_SE, ")"),
         MR_Hevo_Lower_CI = round(Hevo_Lower_CI, 2),
         MR_Hevo_Upper_CI = round(Hevo_Upper_CI, 2),
         MR_Hevo_95_CI = paste0(round(Hevo_Lower_CI, 2), " to ", round(Hevo_Upper_CI, 2)),
         MR_Hevo_Causal_Report_Rate = paste0((Hevo_Pos_Rate * 100), "%")) %>%
  # Drop old names
  select(N,
         Invalid,
         F,
         R2,
         Weighted_Median_Mean_Estimate,
         Weighted_Median_Mean_SE,
         Weighted_Median_95_CI,
         Weighted_Median_Causal_Report_Rate,
         MR_Hevo_Mean_Estimate,
         MR_Hevo_Mean_SE,
         MR_Hevo_95_CI,
         MR_Hevo_Causal_Report_Rate,
         Scenario) %>%
  as_grouped_data(groups = c("Scenario")) %>%
  as_flextable(hide_grouplabel = TRUE, col_keys = c("N",
                                                    "Invalid",
                                                    "F",
                                                    "R2",
                                                    "Weighted_Median_Mean_Estimate",
                                                    "Weighted_Median_Mean_SE",
                                                    "Weighted_Median_95_CI", 
                                                    "Weighted_Median_Causal_Report_Rate",
                                                    "blank_col",
                                                    "MR_Hevo_Mean_Estimate",
                                                    "MR_Hevo_Mean_SE",
                                                    "MR_Hevo_95_CI", 
                                                    "MR_Hevo_Causal_Report_Rate",
                                                    "Scenario")) |>
  # To separate WME/MR-Hevo
  width(j = "blank_col", width = 0.2) |>
  empty_blanks() |>
  span_header() |>
  align(part = "all", align = "center") |>
  align(part = "body", j =c(5, 9),  align = "right") |>
  align(part = "body", j =c(6, 10),  align = "left") |>
  # Italicise N and F
  italic(part = "header", i = 1, j = c(1,3) ) |>
  # Fix name of invalid IVs
  compose(part = "header", i = 1:4, j = 2, value = as_paragraph("Invalid IVs")) |>
  # Fix name of 95% CIs
  compose(part = "header", i = 3, j = c(7, 12), value = as_paragraph("Mean")) |>
  compose(part = "header", i = 4, j = c(7, 12), value = as_paragraph("95% CI")) |>
  # Fix split headers
  compose(part = "header", i = 3, j = c(5,10), value = as_paragraph("Mean Estimate")) |>
  compose(part = "header", i = 4, j = c(5:6,10:11), value = as_paragraph("(Mean SE)")) |>
  merge_h(part = "header", i = 4) |>
  # Italicise/superscript R^2
  compose(part = "header", i = 1, j = 4, value = as_paragraph(as_i("R"), as_sup(as_i(as.character(2))))) |>
  
  hline(part = "header", i = 2, j = c(5:8, 10:13), border = fp_border(color = "black", width = 1)) |>
  #footnote(i = c(1,4), j = c(2, 5), part = "header", value = as_paragraph(c("IV: Instrumental Variable", "SE: Standard Error")))
  # Fix dimensions
  autofit() |>
  #line_spacing(space = 0, part = "header") |>
  line_spacing(space = 1, part = "body") |>
  fit_to_width(max_width = 6.5) |>
  #height_all(height = 40, unit = "mm",part = "all") |>

  # Add footer
  add_footer_lines("CI: Confidence Interval, IV: Instumental Variable, SE: Standard Error. Null Causal Effect (\U03B2 = 0)") |>
  fontsize(8, i = c(1:5), j = c(1:13), part = "header") |>
  fontsize(8, i = c(1:15), j = c(1:13), part = "body") |>
  fontsize(8, i = 1, j = 1, part = "footer")
# set_caption(caption = "Summary of 1000 Simulated Mendelian Randomisation Studies With Null Causal Effect (\U03B2 = 0)",
#             style = "Table Caption",
#             autonum = run_autonum(seq_id = "tab",
#                                   bkm = "no-causal-sim-summ-tib"))

causal_sim_summ_tib_display
#causal_sim_summ_tib


```

\newpage


## Citations Search

A total of 110 abstracts and 54 full texts were screened to identify the 10 studies listed in Table \@ref(tab:citation-search-sample-display); the screening flow diagram is presented in Figure \@ref(fig:screening-flow).


```{r citations-reanalysis-summ-fn, include=FALSE, eval=FALSE}

get_reanalysis_tib <- function(models_in){
  
  output_tib_row <- tibble(N = as.integer(),
                           WME_Av = as.double(),
                           WME_SE = as.double(),
                           Hevo_Av = as.double(),
                           Hevo_SE = as.double()
  )
  
  # Create blank tibble to receive results of Weighted
  # Median Estimator function from MR-Base
  
  results_tib <-  tibble(
                         Study_Author = as.character(),
                         Study_Ref = as.character(),
                         Study_DOI = as.character(),
                         Study_Exposure = as.character(),
                         Study_Outcome = as.character(),
                         Study_Reported_Measure = as.character(),
                         N = as.integer(),
                         WME_est = as.double(),
                         WME_se = as.double(),
                         WME_pval = as.double(),
                         WME_nsnp = as.integer(),
                         Hevo_est = as.double(),
                         Hevo_se = as.double(),
                         Hevo_sd = as.double(),
                         Hevo_est_lower_CI = as.double(),
                         Hevo_est_upper_CI = as.double()
  )
  
  # Convert full tibble to list of 1 tibble per dataset
  model_list <- models_in %>% 
    group_by(Study_DOI) %>%
    group_split()
  
  
  # Should = 10
  n_datasets <- length(model_list)
  
  #Run WME and MR-Hevo for each dataset
  for(dataset in 1:n_datasets){
    
    # Extract study details
    results_tib[dataset, ]$Study_Author <- model_list[[dataset]]$Author[[1]]
    results_tib[dataset, ]$Study_Ref <- model_list[[dataset]]$Study_Ref[[1]]
    results_tib[dataset, ]$Study_DOI <- model_list[[dataset]]$Study_DOI[[1]]
    results_tib[dataset, ]$Study_Exposure <- model_list[[dataset]]$Exposure[[1]]
    results_tib[dataset, ]$Study_Outcome <- model_list[[dataset]]$Outcome[[1]]
    results_tib[dataset, ]$Study_Reported_Measure <- model_list[[dataset]]$Reported_Measure[[1]]

    # Stored as individual vectors for MR-Hevo/RStan - not
    # Tidyverse compatible
    coeff_G_X_vect <- model_list[[dataset]]$Coeff_G_X
    coeff_G_Y_vect <- model_list[[dataset]]$Coeff_G_Y
    coeff_G_X_SE_vect <- model_list[[dataset]]$Coeff_G_X_SE
    coeff_G_Y_SE_vect <- model_list[[dataset]]$Coeff_G_Y_SE


    # N.B. MR-Hevo terminology vs WME paper/other code:
    # alpha = effects of instruments on exposure, i.e. coeff_G_X
    # beta = pleiotropic effects of instruments on outcome, i.e. alpha in WME
    # gamma = effects of instruments on outcome, i.e. coeff_G_Y
    # theta = causal effect X on Y, i.e. b

    # Results from weighted median estimator method
    WME_results <- mr_weighted_median(b_exp = coeff_G_X_vect,
                                      b_out = coeff_G_Y_vect,
                                      se_exp = coeff_G_X_SE_vect,
                                      se_out = coeff_G_Y_SE_vect,
                                      parameters = list(nboot = 1000))

    # Results from MR-Hevo method
    Hevo_results <- run_mrhevo.sstats(alpha_hat = coeff_G_X_vect,
                                      se.alpha_hat = coeff_G_X_SE_vect,
                                      gamma_hat = coeff_G_Y_vect,
                                      se.gamma_hat = coeff_G_Y_SE_vect
    ) %>%
      summary()


    # Extract WME Results
    results_tib[dataset, ]$WME_est <- WME_results$b
    results_tib[dataset, ]$WME_se <- WME_results$se
    results_tib[dataset, ]$WME_pval <- WME_results$pval
    results_tib[dataset, ]$WME_nsnp <- WME_results$nsnp

    # Extract MR-Hevo Results
    results_tib[dataset, ]$Hevo_est <- Hevo_results$summary["theta","mean"]
    results_tib[dataset, ]$Hevo_se <- Hevo_results$summary["theta","se_mean"]
    results_tib[dataset, ]$Hevo_sd <- Hevo_results$summary["theta","sd"]
    results_tib[dataset, ]$Hevo_est_lower_CI <- Hevo_results$summary["theta","2.5%"]
    results_tib[dataset, ]$Hevo_est_upper_CI <- Hevo_results$summary["theta","97.5%"]

    results_tib[dataset, ]$N <- dataset


  }

  results_tib <- results_tib %>%
    mutate(WME_OR = exp(WME_est),
           WME_est_lower_CI = (WME_est - (1.96 * Hevo_se)),
           WME_est_upper_CI = (WME_est + (1.96 * Hevo_se)),
           WME_OR_lower_CI = exp(WME_est_lower_CI),
           WME_OR_upper_CI = exp(WME_est_upper_CI),
           WME_est_causal_detected = (WME_est_lower_CI > 0  | WME_est_upper_CI < 0),
           #WME_OR_causal_detected = (WME_OR_lower_CI > 1  | WME_OR_upper_CI < 1),
           Hevo_OR = exp(Hevo_est),
           Hevo_OR_lower_CI = exp(Hevo_est_lower_CI),
           Hevo_OR_upper_CI = exp(Hevo_est_upper_CI),
           Hevo_OR_causal_detected = (Hevo_OR_lower_CI > 1  | Hevo_OR_upper_CI < 1),
           Hevo_est_causal_detected = (Hevo_est_lower_CI > 0  | Hevo_est_upper_CI < 0)
    )

  return(results_tib)
  
}


```

```{r citations-reanalysis-summ-tib, eval=FALSE}

#citations_reanalysis_summ_tib <-   
  Citations_Instrument_Data %>%
  #slice(1:3108) %>% #trying to select blank rows without this?
  #slice(2282:2730) %>% #xie
  #slice(2731:3108) %>% #Clift
  mutate(across(Coeff_G_X:Coeff_G_Y_SE, ~replace(., . == 0, 10^-100))) %>% 
  get_reanalysis_tib() 

saveRDS(citations_reanalysis_summ_tib, file = here("MSc_Thesis_Split", "Data", "Summary_Tables", "citations_reanalysis_summ_tib.rds"))


```



```{r citation-search-sample-display, include=TRUE, cache=TRUE}
#| tab.id = "citation-search-sample"
#| tab.cap = "Summary of highly-cited two-sample Mendelian randomisation studies reporting a weighted median estimate of casual effect"


citations_search_sample <- read_csv(here("MSc_Thesis_Split", "Data", "Citations_Search", "Citations_Methods_Comparison.csv"),na = "#N/A")

citations_search_sample %>% 
  mutate(
    # fill blanks
    Reported_WME_Lower_CI = if_else(is.na(Reported_WME_Lower_CI), 
                                    (Reported_WME_Causal_Effect - (1.96 * Reported_WME_SE)),
                                    Reported_WME_Lower_CI),
    Reported_WME_Upper_CI = if_else(is.na(Reported_WME_Upper_CI),
                                    (Reported_WME_Causal_Effect + (1.96 * Reported_WME_SE)),
                                    Reported_WME_Upper_CI),
    Overlap_Max_Est_Decimal = replace_na(Overlap_Max_Est_Decimal, 0),
    # round
    across(where(is.numeric), ~round(., 2)),
    # combine effect estimate + CIs
    Effect_Estimate = paste0(Reported_WME_Causal_Effect, " (", Reported_WME_Lower_CI, " to ", Reported_WME_Upper_CI, ")"),
    # combine author/year
    Study = paste0(Author, ", ", Year),
    # convert TRUE/FALSE to Yes/No
    Causality_Reported = if_else(Causality_Reported, "Yes", "No"),
    # convert to %
    Participants_Maximum_Estimated_Overlap = paste0((Overlap_Max_Est_Decimal * 100), "%"),
    # change naming for flextable
    Participants_N_Exposure = Exposure_N_Participants,
    Participants_N_Outcome = Outcome_N_Participants
    
  ) %>% 
  #sort by citation count
  arrange(desc(Citations)) %>% 
  select(Study,
         DOI,
         Citations,
         Association,
         n_Instruments,
         Participants_N_Exposure,
         Participants_N_Outcome,
         Participants_Maximum_Estimated_Overlap,
         Effect_Estimate,
         Causality_Reported,
         p_value
         ) %>% 
  as_flextable() %>% 
  span_header() %>% 
  align(part = "all", align = "center") %>% 
  # Fix column titles
  ## N Instruments
  compose(part = "header", j = 5, value = as_paragraph(as_i("N"), " Instruments")) %>% 
  merge_v(part = "header", j = 5) %>% 
  ## Participants
  compose(part = "header", i = 2, j = 6, value = as_paragraph(as_i("N"))) %>%
  ## Effect
  compose(part = "header", j = 9, value = as_paragraph("Effect Estimate")) %>% 
  merge_v(part = "header", j = 9) %>% 
  ## Causality
  compose(part = "header", j = 10, value = as_paragraph("Causality Reported")) %>% 
  merge_v(part = "header", j = 10) %>% 
  ## p-value 
  compose(part = "header", j = 11, value = as_paragraph(as_i("p"), "-value")) %>% 
  merge_v(part = "header", j = 11) %>% 
  # add lines
  hline(part = "header", i = 1, j = c(6:8), border = fp_border(color = "black", width = 1)) #%>% 
  #hline(part = "header", i = 2, j = c(6:7), border = fp_border(color = "black", width = 1))
  
  
```


```{r citations-reanalysis-summ-display, include=TRUE, cache=TRUE}
#| tab.id = "citations-reanalysis-summ-display",
#| tab.cap = "Reanalysis of 10 highly-cited two-sample MR articles referencing Bowden et al[@bowden_consistent_2016], comparing results of both WME and MR-Hevo causal effect estimation methods",
#| ft.arraystretch = 1

citations_reanalysis_summ_tib <- readRDS(here("MSc_Thesis_Split", "Data", "Summary_Tables", "citations_reanalysis_summ_tib.rds"))

citations_reanalysis_summ_display <- citations_reanalysis_summ_tib %>%
  # Change names for cleaner plotting
  mutate(#Study_Author = paste0(Study_Author, Study_Ref),
         Study_SNPs = WME_nsnp,
         Blank = NA,
         Weighted_Median_Beta = paste0(round(WME_est, 2), " (", round(WME_est_lower_CI, 2), "-", round(WME_est_upper_CI, 2), ")"),
         Weighted_Median_SE = WME_se,
         Weighted_Median_OR = paste0(round(WME_OR, 2), " (", round(WME_OR_lower_CI, 2), "-", round(WME_OR_upper_CI, 2), ")"),
         Weighted_Median_Causality = if_else(WME_est_causal_detected, "Yes", "No"),
         Blank_2 = NA,
         MR_Hevo_Beta = paste0(round(Hevo_est, 2), " (", round(Hevo_est_lower_CI, 2), "-", round(Hevo_est_upper_CI, 2), ")"),
         MR_Hevo_SE = Hevo_se,
         MR_Hevo_OR = paste0(round(Hevo_OR, 2), " (", round(Hevo_OR_lower_CI, 2), "-", round(Hevo_OR_upper_CI, 2), ")"),
         MR_Hevo_Causality = if_else(Hevo_est_causal_detected, "Yes", "No")
         ) %>% 
  # Drop old names
  select(Study_Author,
         #Study_Ref,
         #Study_DOI,
         Study_Exposure,
         Study_Outcome,
         Study_SNPs, 
         Blank,
         Weighted_Median_Beta,
         Weighted_Median_SE,
         Weighted_Median_OR,
         Weighted_Median_Causality,
         Blank_2,
         MR_Hevo_Beta,
         MR_Hevo_SE,
         MR_Hevo_OR,
         MR_Hevo_Causality) %>% 
    mutate(across(where(is.double), signif, 2)) %>% 
  #as_grouped_data(groups = "Study_Author", expand_single = FALSE,) %>%
  flextable(#hide_grouplabel = TRUE, 
    col_keys = c("Study_Author", #"Study_Ref", #"Study_DOI",
                 "Study_Exposure",
                 "Study_Outcome",
                 "Study_SNPs", 
                 "Blank", #5
                 "Weighted_Median_Beta", 
                 "Weighted_Median_SE",
                 "Weighted_Median_OR",
                 "Weighted_Median_Causality",
                 "Blank_2", #10
                 "MR_Hevo_Beta", 
                 "MR_Hevo_SE",
                 "MR_Hevo_OR",
                 "MR_Hevo_Causality")) |>
  # To separate Study Details/WME/MR-Hevo
  width(j = c(5, 10), width = 0.2) |>
  span_header() |>
  align(part = "all", align = "center") |>
  # Fix names blank cols - unsure why not working as above
  compose(part = "header", i = 1:3, j = c(5, 10), value = as_paragraph("")) |>
  # Blank study top row
  compose(part = "header", i = 1, j = c(1), value = as_paragraph("")) |>
  empty_blanks() |>
  # Fix Causality name
  width(j = c(6, 8, 11, 13), width = 20) |>
  compose(part = "header", i = 3, j = c(9, 14), value = as_paragraph("Causality\nReported")) |>
  empty_blanks() |>
  # Fix Beta name
  compose(part = "header", i = 3, j = c(6, 11), value = as_paragraph("\U03B2")) |> #\U0302
  # Separating lines
  hline(part = "header", i = 2, j = c(6:9, 11:14), border = fp_border(color = "black", width = 1)) |>
  # Fix dimensions
  autofit() |>
  line_spacing(space = 1, part = "header") |>
  line_spacing(space = 1, part = "body") |>
  fit_to_width(max_width = 7.5) |>
  #height_all(height = 40, unit = "mm",part = "all") |>
  # Add footer
  add_footer_lines("\U03B2 and OR presented as: estimate (95% CI).\n\U03B2: causal effect estimate, CI: Confidence Interval, OR: Odds Ratio, SE: Standard Error.") |>
  #footnote(i = c(1,4), j = c(2, 5), part = "header", value = as_paragraph(c("IV: Instrumental Variable", "SE: Standard Error")))
  fontsize(8, i = c(1:3), j = c(1:14), part = "header") |>
  fontsize(8, i = c(1:8), j = c(1:14), part = "body") |>
  fontsize(8, i = 1, j = 1, part = "footer")
  # reframe(N = N,
  #         WME_Av = mean(WME_est),
  #         WME_SE = mean(WME_se),
  #         Hevo_Av = mean(Hevo_est),
  #         Hevo_SE = mean(Hevo_se),
  #         Hevo_Est_Causal = Hevo_est_causal_detected
  # ) %>%

citations_reanalysis_summ_display

```

<!-- Table (@\tab:citations-reanalysis-summ-display) reference[@bowden_consistent_2016] -->




