---
title: "9. Appendices"
author: "B233241"
date: "2025-01-07"
output:
  pdf_document:
    latex_engine: xelatex
    dev: cairo_pdf
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile,
                        encoding = encoding,
                        knit_root_dir = rprojroot::find_rstudio_root_file(),
                        output_dir = file.path("../Output/")
                        )})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning=FALSE, 
                      message=FALSE, 
                      error=FALSE)

library(tidyverse)
library(TwoSampleMR)
library(mrhevo)
library(rstan)
library(here)
library(data.table)


```




### Generating Data and Models

The data generating model used was from Appendix 3 of Bowden et al (ref); the relevant section describing their model is reproduced below:

_"..._

\begin{equation} 
U_i = \sum^J_{j=1} \phi_jG_{ij} + \epsilon_i^U
\end{equation}


\begin{equation} 
X_i = \sum^J_{j=1} \gamma_jG_{ij} + U_i + \epsilon_i^X
\end{equation}

\begin{equation} 
Y_i = \sum^J_{j=1} \alpha_jG_{ij} + \beta X_i + U_i + \epsilon_i^Y
\end{equation}

_for participants indexed by $i = 1, . . . , N$, and genetic instruments indexed by $j = 1, . . . , J$._

_The error terms $\epsilon_i^U , \epsilon_i^X$ and $\epsilon_i^Y$ were each drawn independently from standard normal distributions. The genetic effects on the exposure γj are drawn from a uniform distribution between 0.03 and 0.1. Pleiotropic effects $\alpha_j$ and $\phi_j$ were set to zero if the genetic instrument was a valid instrumental variable. Otherwise (with probability 0.1, 0.2, or 0.3):_

_1. In Scenario 1 (balanced pleiotropy, InSIDE satisfied), the $\alpha_j$ parameter was drawn from a uniform distribution between −0.2 and 0.2._

_2. In Scenario 2 (directional pleiotropy, InSIDE satisfied), the $\alpha_j$ parameter was drawn from a uniform distribution between 0 and 0.2._ 

_3. In Scenario 3 (directional pleiotropy, InSIDE not satisfied), the $\phi_j$ parameter was drawn from a uniform distribution between −0.2 and 0.2._


_The causal effect of the exposure on the outcome was either $\beta X = 0$ (null causal effect) or $\beta X = 0.1$ (positive causal effect). A total of 10 000 simulated datasets were generated for sample sizes of N = 10 000 and 20 [sic] participants. Only the summary data, that is genetic associations with the exposure and with the outcome and their standard errors as estimated by univariate regression on the genetic instruments in turn, were used by the analysis methods. In the two-sample setting, data were generated on 2N participants, and genetic associations with the exposure were estimated in the first N participants, and genetic associations with the outcome in the second N participants."_ (ref)

To reproduce this model, code was written in R to generate the relevant participant level data. First, a function (`simulate_MR_data`) was written which included parameters specified by Bowden et al, and also to allow testing of data simulation:



```{r wme-model, echo=TRUE}

# Define function to create data generating model
# Arguments based on Bowden et al
simulate_MR_data <- function(n_participants = as.integer(), 
                             n_instruments = as.integer(),
                             n_datasets = as.integer(),
                             prop_invalid = 0.1,
                             causal_effect = TRUE,
                             balanced_pleio = TRUE,
                             InSIDE_satisfied = TRUE,
                             rand_error = TRUE,              # remove random errors, for testing
                             two_sample = TRUE,              # 1- or 2-sample MR toggle, for testing
                             beta_val = 0.1,                 # size of causal effect
                             allele_freq_min = 0.01,         # frequency of effect allele
                             allele_freq_max = 0.99,
                             gamma_min = 0.03,               # size of pleiotropic effects on exposure
                             gamma_max = 0.1,
                             alpha_min = -0.2,               # size of pleiotropic effects on outcome
                             alpha_max = 0.2,
                             phi_min = -0.2,                 # size of additional pleiotropic effects
                             phi_max = 0.2){                 # when InSIDE not satisfied
  
  
  
  # Initialise blank lists to receive datasets for
  # each of:
  #     U (vector: unmeasured confounding exposures per participant), 
  #     X (vector: exposure:outcome associations estimated per participant) 
  #     Y (vector: gene:outcome association estimated per participant), 
  #     G (Matrices: Genotype data)
  #
  #     gamma (vector: pleiotropic effects of each instrument on exposure)
  #     alpha (vector: pleiotropic effects of each instrument on outcome)
  #     phi (vector: additional pleiotropic effects of each instrument when InSIDE assumption not satisfied)
  U_list <- list()
  X_list <- list()
  Y_list <- list()
  G_X_list <- list()
  G_Y_list <- list()
  
  gamma_list <- list()
  alpha_list <- list()
  phi_list <- list()
  beta_list <- list()
  prop_invalid_list <- list()
  
  
  # --- Assign features common to all datasets ---
  
  beta <- if_else(causal_effect == TRUE, # size of causal effect
                  beta_val,
                  0)
  
  
  
  # --- Create separate datasets ---
  
  # Create N datasets by simulating genotype matrices with
  # 1 row per participant, 1 column per genetic instrument
  # Use these to estimate U, X + Y
  
  for(n in 1:n_datasets){
    
    # Create error terms for U, X + Y per participant,
    # each drawn from standard normal distribution
    # unless random error turned off (for testing)
    
    ifelse(rand_error == TRUE,
           U_epsilon_vect <- rnorm(n = 2 * n_participants),
           U_epsilon_vect <- rep(0, 2 * n_participants))
    
    ifelse(rand_error == TRUE,
           X_epsilon_vect <- rnorm(n = n_participants),
           X_epsilon_vect <- rep(0, n_participants))
    
    ifelse(rand_error == TRUE,
           Y_epsilon_vect <- rnorm(n = n_participants),
           Y_epsilon_vect <- rep(0, n_participants))
    
    
    # --- Create matrix of genotypes --- 
    
    # 0 = reference, i.e. zero effect alleles, 
    # 1 = 1 effect allele, 2 = 2 effect alleles 
    
    
    # Probability of effect allele set per dataset  
    # for each instrument, default value set at  
    # random between 0.01-0.99 (i.e. both effect +
    # reference are common alleles)
    allele_freq_vect <- runif(n = n_instruments,
                              min = allele_freq_min,
                              max = allele_freq_max)
    
    
    
    # Assign genotypes by sampling from binomial distribution
    # twice (as two alleles) per participant with probability
    # equal to frequency of effect allele
    # Create twice as many genotypes as participants in sample
    # to simulate 2 sample MR, i.e. first half used to estimate
    # Gene:Exposure, second half used to estimate Gene:Outcome
    
    # Matrix where columns are instruments, rows are participants
    # Values 0, 1 or 2
    
    
    G_mat <- matrix(rbinom(n = 2 * n_participants * n_instruments,
                           size = 2,
                           prob = rep(allele_freq_vect, 2 * n_participants)),
                    nrow = 2 * n_participants,
                    ncol = n_instruments,
                    byrow = TRUE)
    
    
    # --- Set characteristics for each genetic instrument ---
    
    # Set which instruments invalid
    invalid_instrument_vect <- rbinom(n = n_instruments,
                                      size = 1, 
                                      prob = prop_invalid)
    
    
    # Set genetic effects of each instrument on the exposure,
    # drawn from uniform distribution, min/max as per Bowden 
    # et al
    gamma_vect <- runif(n = n_instruments,
                        min = gamma_min,
                        max = gamma_max)
    
    
    # Set pleiotropic effects on outcome, Scenarios and 
    # min/max from Bowden et al
    alpha_vect <- double() # Pleiotropic effects of instruments on outcome
    phi_vect <- double() # Pleiotropic effects of confounders on outcome
    
    for(j in 1:n_instruments){
      ifelse(invalid_instrument_vect[j] == FALSE,
             alpha_vect[j] <- 0,
             ifelse(balanced_pleio == TRUE,
                    alpha_vect[j] <- runif(n = n_instruments,
                                           min = alpha_min,
                                           max = alpha_max),
                    alpha_vect[j] <- runif(n = n_instruments,
                                           min = 0,
                                           max = alpha_max)
             )
      )
      
      # Assign default phi = 0 unless unbalanced pleiotropy & 
      # InSIDE assumption not satisfied & genetic instrument invalid
      if(balanced_pleio == FALSE & InSIDE_satisfied == FALSE){
        ifelse(invalid_instrument_vect[j] == FALSE,
               phi_vect[j] <- 0,
               phi_vect[j] <- runif(n = 1,
                                    min = phi_min,
                                    max = phi_max)
        )
        
      }
      else{
        phi_vect[j] <- 0
      }
    }
    
    
    # --- Combine Gene matrix/parameters to recreate model --- 
    
    # Create vectors of estimates for U, X and Y per individual,
    # i.e. Ui, Xi and Yi. Uses matrix inner product operator " %*%" 
    # https://stackoverflow.com/questions/22060515/the-r-operator 
    # http://matrixmultiplication.xyz/
    
    
    Ui_vect <-  G_mat %*% phi_vect + U_epsilon_vect
    
    Xi_vect <-  G_mat[1:n_participants, ] %*% gamma_vect + 
      Ui_vect[1:n_participants, ] + 
      X_epsilon_vect
    
    Yi_vect <-  G_mat[(n_participants+1):(2*n_participants), ] %*% alpha_vect + 
      beta * Xi_vect + 
      Ui_vect[(n_participants+1):(2*n_participants), ] + 
      Y_epsilon_vect
    
    
    # Add vectors of estimates from this dataset to lists of 
    # estimates from all datasets
    U_list[[n]] <- Ui_vect
    
    X_list[[n]] <- Xi_vect
    
    Y_list[[n]] <- Yi_vect
    
    G_X_list[[n]] <- G_mat[1:n_participants, ]
    
    # Default is to estimate gene:outcome coefficient from different sample
    # to gene:exposure coefficient (i.e. simulating 2-sample MR)
    # two_sample == FALSE toggles to single sample for testing simulation
    ifelse(two_sample == TRUE,
           G_Y_list[[n]] <- G_mat[(n_participants+1):(2*n_participants), ],
           G_Y_list[[n]] <- G_mat[1:n_participants, ])
    
    
    alpha_list[[n]] <- alpha_vect
    
    gamma_list[[n]] <- gamma_vect
    
    phi_list[[n]] <- phi_vect
    
    beta_list[[n]] <- beta
    
    prop_invalid_list[[n]] <- prop_invalid
    
    
  }
  
  #     U (vector: unmeasured confounding exposures per participant), 
  #     X (vector: exposure:outcome associations estimated per participant) 
  #     Y (vector: gene:outcome association estimated per participant) 
  
  
  # --- Combine all outputs to return --- 
  combined_list <- list(U = U_list,         # Estimates 
                        X = X_list, 
                        Y = Y_list,
                        G_X = G_X_list,     # Genotypes of 1st sample
                        G_Y = G_Y_list,     # Genotypes of 2nd sample
                        alpha = alpha_list, # Actual values for validating simulation
                        gamma = gamma_list,
                        phi = phi_list,
                        beta = beta_list,
                        prop_invalid = prop_invalid_list
  )
  
  return(combined_list)
  
}
```


A function (`extract_models`) was then written to create linear models from each dataset generated as per Bowden et al:


```{r extract-models}


# Create plotting tibble with Mean/SD X + Y grouped by
# Dataset + instrument
extract_models <- function(sim){
  
  output_list <- list()
  
  # Create linear models per dataset to get coefficients
  # for gene:exposure association (coeff_G_X) and gene:outcome
  # association (coeff_G_Y)
  for(dataset in 1:length(sim$X)){
    
    X <- sim$X[[dataset]]
    Y <- sim$Y[[dataset]]
    Instruments_X <- sim$G_X[[dataset]]
    Instruments_Y <- sim$G_Y[[dataset]]
    
    alpha <- sim$alpha[[dataset]]
    gamma <- sim$gamma[[dataset]]
    phi <- sim$phi[[dataset]]
    beta <- sim$beta[[dataset]]
    prop_invalid <- sim$prop_invalid[[dataset]]
    
    
    # Model for gene:exposure
    X_lm <- lm(X ~ 0 + Instruments_X)
    coeff_G_X_vect <- coef(summary(X_lm))[1:(ncol(Instruments_X)), 1]
    SE_coeff_G_X_vect <- coef(summary(X_lm))[1:(ncol(Instruments_X)), 2]
    
    # Model for gene:outcome
    Y_lm <- lm(Y ~ 0 + Instruments_Y)
    coeff_G_Y_vect <- coef(summary(Y_lm))[1:(ncol(Instruments_Y)), 1] 
    SE_coeff_G_Y_vect <- coef(summary(Y_lm))[1:(ncol(Instruments_Y)), 2]
    
    output_list[[dataset]] <- as_tibble(list(dataset = dataset,
                                             Instrument = c(1:ncol(Instruments_X)),
                                             coeff_G_X = coeff_G_X_vect,
                                             coeff_G_X_SE = SE_coeff_G_X_vect,
                                             gamma = gamma,
                                             coeff_G_Y = coeff_G_Y_vect,
                                             coeff_G_Y_SE = SE_coeff_G_Y_vect,
                                             alpha = alpha,
                                             phi = phi,
                                             beta = beta,
                                             prop_invalid = prop_invalid),
                                        .name_repair = "unique")
  }
  
  return(output_list)
  
}  


```




```{r edit-summary-MR-tib-row-fn, warning=FALSE}


edit_get_summary_MR_tib_row <- function(n_participants = 1000,          # update to 10,000 for final
                                   n_instruments = 25,
                                   n_datasets = 100,               # update to 10,000 for final
                                   prop_invalid = 0.1,
                                   causal_effect = TRUE,
                                   balanced_pleio = TRUE,
                                   InSIDE_satisfied = TRUE,
                                   rand_error = TRUE,
                                   two_sample = TRUE, 
                                   beta_val = 0.1,
                                   allele_freq_min = 0.01,
                                   allele_freq_max = 0.99,
                                   gamma_min = 0.03,               
                                   gamma_max = 0.1,
                                   alpha_min = -0.2,               
                                   alpha_max = 0.2,
                                   phi_min = -0.2,                 
                                   phi_max = 0.2){        
  
  
  # Generate data as per Bowden models with helper function 
  dataset_list <- simulate_MR_data(n_participants = n_participants,          
                                   n_instruments = n_instruments,
                                   n_datasets = n_datasets,               
                                   prop_invalid = prop_invalid,
                                   causal_effect = causal_effect,
                                   balanced_pleio = balanced_pleio,
                                   InSIDE_satisfied = InSIDE_satisfied,
                                   rand_error = rand_error,     
                                   two_sample = two_sample,
                                   beta_val = beta_val,                 
                                   allele_freq_min = allele_freq_min,         
                                   allele_freq_max = allele_freq_max,
                                   gamma_min = gamma_min,                
                                   gamma_max = gamma_max,
                                   alpha_min = alpha_min,               
                                   alpha_max = alpha_max,
                                   phi_min = phi_min,                 
                                   phi_max = phi_max)
  
  
  # Use data to create model per dataset
  model_list <- extract_models(dataset_list)
  
  
  # Create output tibble in same format as Table 2/3 from
  # Bowden et al
  output_tib_row <- tibble(N = as.integer(),
                           Prop_Invalid = as.double(),
                           F_stat = as.double(),
                           R2_stat = as.double(),
                           WME_Av = as.double(),
                           WME_SE = as.double(),
                           WME_Pos_Rate = as.double(),
                           Hevo_Av = as.double(),
                           Hevo_SE = as.double(),
                           Hevo_Pos_Rate = as.double())
  
  output_tib_row$N <-  n_datasets
  
  # Create blank tibble to receive results of Weighted
  # Median Estimator function from MR-Base
  
  results_tib <-  tibble(WME_est = as.double(),
                         WME_se = as.double(),
                         WME_pval = as.double(),
                         WME_Q = as.double(),
                         WME_Q_df = as.double(),
                         WME_Q_pval = as.double(),
                         WME_nsnp = as.integer(),
                         Hevo_est = as.double(),
                         Hevo_se = as.double(),
                         Hevo_sd = as.double(),
                         Hevo_2.5 = as.double(),
                         Hevo_25 = as.double(),
                         Hevo_50 = as.double(),
                         Hevo_75 = as.double(),
                         Hevo_97.5 = as.double(),
                         Hevo_n_eff = as.double(),
                         Hevo_n_Rhat = as.double(),
                         Hevo_z_stat = as.double(),
                         Hevo_pval = as.double(),
                         Hevo_causal_detected = as.logical()
  )

  
  # Run WME and MR-Hevo for each dataset 
  for(dataset in 1:n_datasets){
    
    # Stored as individual vectors for MR-Hevo/RStan - not
    # Tidyverse compatible
    coeff_G_X_vect <- model_list[[dataset]]$coeff_G_X
    coeff_G_Y_vect <- model_list[[dataset]]$coeff_G_Y
    coeff_G_X_SE_vect <- model_list[[dataset]]$coeff_G_X_SE
    coeff_G_Y_SE_vect <- model_list[[dataset]]$coeff_G_Y_SE
    
    
    # N.B. MR-Hevo terminology vs WME paper/other code:
    # alpha = effects of instruments on exposure, i.e. coeff_G_X
    # beta = pleiotropic effects of instruments on outcome, i.e. alpha in WME
    # gamma = effects of instruments on outcome, i.e. coeff_G_Y
    # theta = causal effect X on Y, i.e. b
    
    # Results from weighted median estimator method
    WME_results <- mr_weighted_median(b_exp = coeff_G_X_vect,
                                      b_out = coeff_G_Y_vect,
                                      se_exp = coeff_G_X_SE_vect,
                                      se_out = coeff_G_Y_SE_vect,
                                      parameters = list(nboot = 1000))
    
    # Results from MR-Hevo method
    Hevo_results<- run_mrhevo.sstats(alpha_hat = coeff_G_X_vect,
                                     se.alpha_hat = coeff_G_X_SE_vect,
                                     gamma_hat = coeff_G_Y_vect,
                                     se.gamma_hat = coeff_G_Y_SE_vect#,
                                     #fraction_pleio = 0.2,
                                     #priorsd_theta = 1,
                                     #model.dir = here("MSc_Thesis_Split", "Script", "Hevo")
                                     ) %>%
      summary()

    
    
    results_tib[dataset, ]$WME_est <- WME_results$b
    results_tib[dataset, ]$WME_se <- WME_results$se
    results_tib[dataset, ]$WME_pval <- WME_results$pval
    results_tib[dataset, ]$WME_Q <- WME_results$Q
    results_tib[dataset, ]$WME_Q_df <- WME_results$Q_df
    results_tib[dataset, ]$WME_Q_pval <- WME_results$Q_pval
    results_tib[dataset, ]$WME_nsnp <- WME_results$nsnp
    
    # Extract MR-Hevo Results
    results_tib[dataset, ]$Hevo_est <- Hevo_results$summary["theta","mean"]
    results_tib[dataset, ]$Hevo_se <- Hevo_results$summary["theta","se_mean"]
    results_tib[dataset, ]$Hevo_sd <- Hevo_results$summary["theta","sd"]
    results_tib[dataset, ]$Hevo_2.5 <- Hevo_results$summary["theta","2.5%"]
    results_tib[dataset, ]$Hevo_25 <- Hevo_results$summary["theta","25%"]
    results_tib[dataset, ]$Hevo_50 <- Hevo_results$summary["theta","50%"]
    results_tib[dataset, ]$Hevo_75 <- Hevo_results$summary["theta","75%"]
    results_tib[dataset, ]$Hevo_97.5 <- Hevo_results$summary["theta","97.5%"]
    results_tib[dataset, ]$Hevo_n_eff <- Hevo_results$summary["theta","n_eff"]
    results_tib[dataset, ]$Hevo_n_Rhat <- Hevo_results$summary["theta","Rhat"]
    
    
  }

  
  results_tib <- results_tib %>% 
    mutate(Hevo_causal_detected = !(Hevo_2.5 < 0  & Hevo_97.5 > 0))
  
  output_tib_row <- results_tib %>% 
    summarise(N = n_datasets,
              Prop_Invalid = prop_invalid,
              F_Stat = NA,
              R2_stat = NA,
              WME_Av = mean(WME_est),
              WME_SE = mean(WME_se),
              WME_Pos_Rate = length(WME_pval[WME_pval < 0.05]) / n_datasets,
              Hevo_Av = mean(Hevo_est),
              Hevo_SE = mean(Hevo_se),
              Hevo_Pos_Rate = sum(Hevo_causal_detected) / n_datasets#[Hevo_pval < (0.05/n_datasets) ] ) / n_datasets
              #Hevo_Pos_Rate = length(Hevo_pval[Hevo_pval < (0.05/n_datasets) ] ) / n_datasets
    ) %>% 
    mutate(across(where(is.double), round, 3))
  
  #return(output_tib_row)
  return(results_tib)
  #return(coeff_G_X_vect)
}

```


```{r edit-test}

# Standard set-up for RStan
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE, save_dso = TRUE)

# Run local MR-Hevo functions file
# Changes to suppress messages/prevent model recompiling
source(here("MSc_Thesis_Split", "Script", "Hevo", "functions.mrhevo.R"))

# Compile model for MR-Hevo
mr.stanmodel <- stan_model(file= here("MSc_Thesis_Split", "Script", "Hevo", "MRHevo_summarystats.stan"),
                                 model_name="MRHevo.summarystats", verbose=FALSE)

set.seed(12345)
test_edit <-  edit_get_summary_MR_tib_row(n_participants = 10000,
                                            n_instruments = 25,
                                            n_datasets = 1,
                                            prop_invalid = 0,
                                            beta_val = 0.1,
                                            causal_effect = TRUE,
                                            rand_error = FALSE, 
                                            two_sample = FALSE,
                                            balanced_pleio = TRUE,
                                            InSIDE_satisfied = TRUE)

test_edit


```


```{r pkg-summary-MR-tib-row-fn, warning=FALSE, include=FALSE}


#source(here("MSc_Thesis_Split", "Script", "Hevo", "functions.mrhevo.R"))
#source(here("MSc_Thesis_Split", "Script", "Hevo", "stanmodels.R"))
#source(here("MSc_Thesis_Split", "Script", "Hevo", "MRHevo_summarystats.stan"))

# Standard set-up for RStan
#options(mc.cores = parallel::detectCores())
#rstan_options(auto_write = TRUE, save_dso = TRUE)

# Run local MR-Hevo functions file
# Changes to suppress messages/prevent model recompiling
#source(here("functions.mrhevo.R"))

# Compile model for MR-Hevo
#mr.stanmodel <- stan_model(file= here("MSc_Thesis_Split", "Script", "Hevo", "MRHevo_summarystats.stan"),
#                                 model_name="MRHevo.summarystats", verbose=FALSE)
# --- Need to comment out model.dir in Hevo_results --- #


pkg_get_summary_MR_tib_row <- function(n_participants = 1000,          # update to 10,000 for final
                                       n_instruments = 25,
                                       n_datasets = 100,               # update to 10,000 for final
                                       prop_invalid = 0.1,
                                       causal_effect = TRUE,
                                       balanced_pleio = TRUE,
                                       InSIDE_satisfied = TRUE,
                                       rand_error = TRUE,
                                       two_sample = TRUE, 
                                       beta_val = 0.1,
                                       allele_freq_min = 0.01,
                                       allele_freq_max = 0.99,
                                       gamma_min = 0.03,               
                                       gamma_max = 0.1,
                                       alpha_min = -0.2,               
                                       alpha_max = 0.2,
                                       phi_min = -0.2,                 
                                       phi_max = 0.2){        
  
  
  # Generate data as per Bowden models with helper function 
  dataset_list <- simulate_MR_data(n_participants = n_participants,          
                                   n_instruments = n_instruments,
                                   n_datasets = n_datasets,               
                                   prop_invalid = prop_invalid,
                                   causal_effect = causal_effect,
                                   balanced_pleio = balanced_pleio,
                                   InSIDE_satisfied = InSIDE_satisfied,
                                   rand_error = rand_error,     
                                   two_sample = two_sample,
                                   beta_val = beta_val,                 
                                   allele_freq_min = allele_freq_min,         
                                   allele_freq_max = allele_freq_max,
                                   gamma_min = gamma_min,                
                                   gamma_max = gamma_max,
                                   alpha_min = alpha_min,               
                                   alpha_max = alpha_max,
                                   phi_min = phi_min,                 
                                   phi_max = phi_max)
  
  
  # Use data to create model per dataset
  model_list <- extract_models(dataset_list)
  
  
  # Create output tibble in same format as Table 2/3 from
  # Bowden et al
  output_tib_row <- tibble(N = as.integer(),
                           Prop_Invalid = as.double(),
                           F_stat = as.double(),
                           R2_stat = as.double(),
                           WME_Av = as.double(),
                           WME_SE = as.double(),
                           WME_Pos_Rate = as.double(),
                           Hevo_Av = as.double(),
                           Hevo_SE = as.double(),
                           Hevo_Pos_Rate = as.double())
  
  output_tib_row$N <-  n_datasets
  
  # Create blank tibble to receive results of Weighted
  # Median Estimator function from MR-Base
  
  results_tib <-  tibble(WME_est = as.double(),
                         WME_se = as.double(),
                         WME_pval = as.double(),
                         WME_Q = as.double(),
                         WME_Q_df = as.double(),
                         WME_Q_pval = as.double(),
                         WME_nsnp = as.integer(),
                         Hevo_est = as.double(),
                         Hevo_se = as.double(),
                         Hevo_sd = as.double(),
                         Hevo_2.5 = as.double(),
                         Hevo_25 = as.double(),
                         Hevo_50 = as.double(),
                         Hevo_75 = as.double(),
                         Hevo_97.5 = as.double(),
                         Hevo_n_eff = as.double(),
                         Hevo_n_Rhat = as.double(),
                         Hevo_z_stat = as.double(),
                         Hevo_pval = as.double(),
                         Hevo_causal_detected = as.logical()
  )

  
  # Run WME and MR-Hevo for each dataset 
  for(dataset in 1:n_datasets){
    
    # Stored as individual vectors for MR-Hevo/RStan - not
    # Tidyverse compatible
    coeff_G_X_vect <- model_list[[dataset]]$coeff_G_X
    coeff_G_Y_vect <- model_list[[dataset]]$coeff_G_Y
    coeff_G_X_SE_vect <- model_list[[dataset]]$coeff_G_X_SE
    coeff_G_Y_SE_vect <- model_list[[dataset]]$coeff_G_Y_SE
    
    
    # N.B. MR-Hevo terminology vs WME paper/other code:
    # alpha = effects of instruments on exposure, i.e. coeff_G_X
    # beta = pleiotropic effects of instruments on outcome, i.e. alpha in WME
    # gamma = effects of instruments on outcome, i.e. coeff_G_Y
    # theta = causal effect X on Y, i.e. b
    
    # Results from weighted median estimator method
    WME_results <- mr_weighted_median(b_exp = coeff_G_X_vect,
                                      b_out = coeff_G_Y_vect,
                                      se_exp = coeff_G_X_SE_vect,
                                      se_out = coeff_G_Y_SE_vect,
                                      parameters = list(nboot = 100000))
    
    # Results from MR-Hevo method
    Hevo_results<- mrhevo::run_mrhevo.sstats(alpha_hat = coeff_G_X_vect,
                                     se.alpha_hat = coeff_G_X_SE_vect,
                                     gamma_hat = coeff_G_Y_vect,
                                     se.gamma_hat = coeff_G_Y_SE_vect,
                                     #fraction_pleio = 0.2,
                                     #priorsd_theta = 1,
                                     model.dir = here("MSc_Thesis_Split", "Script", "Hevo")
                                     ) %>%
      summary()

    
    
    results_tib[dataset, ]$WME_est <- WME_results$b
    results_tib[dataset, ]$WME_se <- WME_results$se
    results_tib[dataset, ]$WME_pval <- WME_results$pval
    results_tib[dataset, ]$WME_Q <- WME_results$Q
    results_tib[dataset, ]$WME_Q_df <- WME_results$Q_df
    results_tib[dataset, ]$WME_Q_pval <- WME_results$Q_pval
    results_tib[dataset, ]$WME_nsnp <- WME_results$nsnp
    
    # Extract MR-Hevo Results
    results_tib[dataset, ]$Hevo_est <- Hevo_results$summary["theta","mean"]
    results_tib[dataset, ]$Hevo_se <- Hevo_results$summary["theta","se_mean"]
    results_tib[dataset, ]$Hevo_sd <- Hevo_results$summary["theta","sd"]
    results_tib[dataset, ]$Hevo_2.5 <- Hevo_results$summary["theta","2.5%"]
    results_tib[dataset, ]$Hevo_25 <- Hevo_results$summary["theta","25%"]
    results_tib[dataset, ]$Hevo_50 <- Hevo_results$summary["theta","50%"]
    results_tib[dataset, ]$Hevo_75 <- Hevo_results$summary["theta","75%"]
    results_tib[dataset, ]$Hevo_97.5 <- Hevo_results$summary["theta","97.5%"]
    results_tib[dataset, ]$Hevo_n_eff <- Hevo_results$summary["theta","n_eff"]
    results_tib[dataset, ]$Hevo_n_Rhat <- Hevo_results$summary["theta","Rhat"]
    
    
  }
  
  # Calculate MR-Hevo p-values from CI, based on https://doi.org/10.1136/bmj.d2304:
  #
  # If the upper and lower limits of a 95% CI are u and l respectively:
  # - 1) calculate the standard error: SE = (u − l)/(2×1.96)
  # - 2) calculate the test statistic: z = Est/SE
  # - 3) calculate the P value2: P = exp(−0.717×z − 0.416×z2).
  
  results_tib <- results_tib %>% 
    mutate(Hevo_z_stat = Hevo_est/Hevo_se,
           Hevo_pval = exp(-0.717 * Hevo_z_stat - 0.416* Hevo_z_stat^2 ),
           Hevo_causal_detected = !(Hevo_2.5 < 0  & Hevo_97.5 > 0))
           #Hevo_p_threshold = 0.05 / n_datasets)
  
  output_tib_row <- results_tib %>% 
    summarise(N = n_datasets,
              Prop_Invalid = prop_invalid,
              F_Stat = NA,
              R2_stat = NA,
              WME_Av = mean(WME_est),
              WME_SE = mean(WME_se),
              WME_Pos_Rate = length(WME_pval[WME_pval < 0.05]) / n_datasets,
              Hevo_Av = mean(Hevo_est),
              Hevo_SE = mean(Hevo_se),
              Hevo_Pos_Rate = length(Hevo_causal_detected)#[Hevo_pval < (0.05/n_datasets) ] ) / n_datasets
              #Hevo_Pos_Rate = length(Hevo_pval[Hevo_pval < (0.05/n_datasets) ] ) / n_datasets
    ) %>% 
    mutate(across(where(is.double), round, 3))
  
  #return(output_tib_row)
  return(results_tib)
  
}

```


```{r pkg-test, include=FALSE}



#set.seed(14101583)

library(mrhevo)
set.seed(1701)
test_pkg <-  pkg_get_summary_MR_tib_row(n_participants = 10000,
                                            n_instruments = 25,
                                            n_datasets = 1,
                                            prop_invalid = 0.1,
                                            beta_val = 10000,
                                            causal_effect = TRUE,
                                            rand_error = FALSE, 
                                            two_sample = FALSE,
                                            balanced_pleio = TRUE,
                                            InSIDE_satisfied = TRUE)


print(test_pkg)

```